A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-01-22 sections/Metatheory.tex
Decl. subtyping figure; Vdash symbol; transfer rules

@@

@@

\begin{theorem}[Soundness]
If \emph{wf }$\Gm$ and $\Gm \redto \cdot$, then $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\cdot$.
\end{theorem}

@@

\begin{theorem}[Completeness]
If \emph{wf }$\Gm$ and $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\cdot$, then $\Gm \redto \cdot$.
\end{theorem}

\begin{lemma}[Prune Transfer for Instantiation]
If $(\Gm \vdash \al \le A \to B) \sto (\Om \vdash C \le A_1 \to B_1)$ and $\Om \vdash_d C \le A_1 \to B_1$,
then $\al\notin FV(A) \cup FV(B)$.
\end{lemma}

\subsection{Decidability}
\begin{theorem}[Decidability]
For all well-formed $\Gm$, it is decidable whether $\Gm\redto\cdot$ or not.
\end{theorem}

@@

\begin{definition}[Worklist Measures]
\[|\Gm|_e = \sum_{e\Lto A\in\Gm}|e|_e + \sum_{e\To_a \jg\in\Gm}|e|_e + \sum_{\appInfAlg{A}{e}\in\Gm}|e|_e\]
\[|\Gm|_\Leftrightarrow = \#_{e\Lto A\in\Gm}
    + 2\cdot\#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\]
\[|\Gm|_\forall = \sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
    \sum_{A\le B\in\Gm} |A|_\forall + |B|_\forall\]
\[|\Gm|_\to = \sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
    \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\]
\[|\Gm|_{\al} = \#_{\al\in\Gm}\]
\end{definition}

>>>>

The transfer rules are shown in Figure~\ref{fig:trans}.
The rules recursively replace the first existential variable to a well-scoped mono-type,
so the algorithmic worklist will finally become a declarative one.
In order to maintain well-formedness,
the substitution should affect all the judgments and term variable bindings in the worklist.

@@

\begin{figure}[t]
\framebox{$\Om \rto \Om'$} $\Om$ reduces to $\Om'$.
\begin{gather*}
\begin{aligned}
\Om,a &\rto \Om \qquad \Om,\al \rto \Om \qquad \Om,x:A \rto \Om\\
\Om\Vdash A\le B &\rto \Om \text{ when } \Om\vdash A\le B\\
\Om\Vdash e\Lto A &\rto \Om \text{ when } \Om\vdash e\Lto A\\
\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg \text{ when } \Om\vdash e\To A\\
\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg \text{ when } \Om\vdash \appInf{A}{e}{C}\\
\end{aligned}
\end{gather*}
\caption{Algorithmic Worklist Typing Rules}
\label{fig:alg:typing}
\end{figure}

@@

\begin{theorem}[Soundness]
If \emph{wf }$\Gm$ and $\Gm \redto \nil$, then $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$.
\end{theorem}

@@

\begin{theorem}[Completeness]
If \emph{wf }$\Gm$ and $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$, then $\Gm \redto \nil$.
\end{theorem}

\begin{lemma}[Prune Transfer for Instantiation]
If $(\Gm \Vdash \al \le A \to B) \sto (\Om \Vdash C \le A_1 \to B_1)$ and $\Om \vdash C \le A_1 \to B_1$,
then $\al\notin FV(A) \cup FV(B)$.
\end{lemma}

\subsection{Decidability}
\begin{theorem}[Decidability]
For all well-formed $\Gm$, it is decidable whether $\Gm\redto\nil$ or not.
\end{theorem}

@@

\begin{definition}[Worklist Measures]
\[|\Gm|_e = \sum_{e\Lto A\in\Gm}|e|_e + \sum_{e\To_a \jg\in\Gm}|e|_e + \sum_{\appInfAlg{A}{e}\in\Gm}|e|_e\]
\[|\Gm|_\Leftrightarrow = \#_{e\Lto A\in\Gm}
    + 2\nil\#_{e\To_a \jg\in\Gm} + 3\nil\#_{\appInfAlg{A}{e}\in\Gm}\]
\[|\Gm|_\forall = \sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
    \sum_{A\le B\in\Gm} |A|_\forall + |B|_\forall\]
\[|\Gm|_\to = \sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
    \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\]
\[|\Gm|_{\al} = \#_{\al\in\Gm}\]
\end{definition}

### 2019-01-22 sections/Overview.tex
Decl. subtyping figure; Vdash symbol; transfer rules

>>>>

\begin{figure}[t]
\[
\begin{array}{l@{\qquad}lcl}
\text{Type variables}\qquad&a, b\\[3mm]
\text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \forall a. A \mid A\to B\\
\text{Monotypes}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
\text{Contexts}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a\\
%\text{Judgments}\qquad&\exps &::=&\quad \cdot \mid A \le B : \exps
\end{array}
\]
\caption{Syntax of Declarative System}\label{fig:decl:syntax}
\end{figure}

The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.

\begin{figure}[t]
\centering \framebox{$\Om \vdash A$}
\begin{gather*}
\inferrule*[right=$\mathtt{wf_d unit}$]
    {~}{\Om\vdash 1}
\qquad
\inferrule*[right=$\mathtt{wf_d var}$]
    {a\in\Om}{\Om\vdash a}
\qquad
\inferrule*[right=$\mathtt{wf_d{\to}}$]
    {\Om\vdash A\quad \Om\vdash B}
    {\Om\vdash A\to B}
\qquad
\inferrule*[right=$\mathtt{wf_d\forall}$]
    {\Om, a\vdash A}
    {\Om\vdash \forall a. A}
\end{gather*}

\centering \framebox{$\Om \vdash A \le B$}
\begin{gather*}
\inferrule*[right=$\mathtt{{\le}Var}$]
    {a\in\Om}{\Om\vdash a\le a}
\qquad
\inferrule*[right=$\mathtt{{\le}Unit}$]
    {~}{\Om \vdash 1 \le 1}
\qquad
\inferrule*[right=$\mathtt{{\le}{\to}}$]
    {\Om \vdash B_1 \le A_1 \quad \Om \vdash A_2 \le B_2}
    {\Om\vdash A_1\to A_2 \le B_1\to B_2}
\\
\inferrule*[right=$\mathtt{{\le}\forall L}$]
    {\Om\vdash \tau \quad \Om\vdash [\tau/a] A \le B}
    {\Om\vdash \forall a. A \le B}
\qquad
\inferrule*[right=$\mathtt{{\le}\forall R}$]
    {\Om, a\vdash A\le B}
    {\Om\vdash A \le \forall a. B}
\end{gather*}
\caption{Well-formedness of Declarative Types and Declarative Subtyping}\label{fig:decl:wf_sub}
\end{figure}

In Figure~\ref{fig:decl:wf_sub},
we give the well-formedness and subtyping relation for the declarative system.
